---
title: "R-markdown Practice"
author: "Amjad"
date: "2024-12-09"
output:
  pdf_document: default
  html_document: default
---


# Day 1
## How to activate repositories------ it will be for certain projects. Should do it in every new project
## setRepositories()
## How to install packages------
## install.packages("tidyverse", dependencies = TRUE)

## Activate Package

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
data(iris)

library(plotly)
fig <-  plot_ly(x = c(0,1, 2), y = c(2, 1, 3), type = 'bar') %>%
  layout(title = 'A Figure Displayed with print(fig)',
         plot_bgcolor='#e5ecf6', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'), 
         yaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2, 
           gridcolor = 'ffff'))

print(fig)

ggplot(iris, aes(Species, Sepal.Length, Sepal.Width, col=Species))+geom_boxplot()

```

# Day 2
```{r}
library(ggplot2)
# code to generate any type of graph
#     ggplot(data = df, aes(x=names, y=height)) + geom_name of particular graph



# Data Import 
data(iris)
View(iris)


# SCATTER PLOT
# import or activate packages

library(ggplot2)
ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width))+geom_point()

# Adding colours

ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width, col=Species))+geom_point()

# Change based on size

ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width, size=Species))+geom_point()

# change shap

ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width,
                              shape=Species))+geom_point()

# multiple aesthetics

ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width, shape=Species, col=Species))+geom_point()

# try to use shape or color do not use both

ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width, col=Species))+geom_point(size=2)



# plot on mpg

data(mpg)
ggplot(data = mpg, aes(x = cty, y = hwy))+geom_point()
ggplot(data = mpg, aes(x = cty, y = hwy, colour =manufacturer))+geom_point()

# to veiw zoomed file Cntrl+shift+6


# to save 

P1 <- ggplot(data = mpg, aes(x = cty, y = hwy, colour =manufacturer))+geom_point()
P1

ggsave("mpg_plot.pdf", plot = P1, width = 6, height = 5, units = "in")


ggsave("mpg_plot.png", plot = P1, width = 6, height = 5, units = "in", dpi = 600)




# Tidyverse

library(tidyverse)


x <- c(1, 2, 3, 4, 5)
y <- c(3, 7, 8, 9, 12)

plot(x, y)


data(women)
P1 <- ggplot(data = women, mapping = aes(x = height, y = weight, colour = weight)) + geom_point()
ggsave("women_plot.pdf", plot = P1, width = 6, height = 5, units = "in")
```


# Day 3
```{r}
data("iris")
df <- iris


# How to save data frames to excel format

library(writexl)
write_xlsx(df, "iris.xlsx")


# import excel file

library(readxl)



df <- read_excel("iris.xlsx")




# plotting in ggplot2

library(ggplot2)
ggplot(data = df, mapping = aes(x = Sepal.Length, y=Sepal.Width, colour = Species))+
  geom_point() + facet_wrap(~Species) + theme_bw()+ theme(panel.grid = element_blank()) +  # removing grid
  labs(title = "Comparison Plot", x = "Hight", y = "Weight", colour = "Legends") + 
  theme(strip.background = element_rect(fill = "lightblue")) + #Strip colour
  theme(strip.text = element_text(colour = "red"))  # strip text colour
  




# How to change facet_labels?
# How to colour facet_labels rectangle?
# How to remove grid lines in the plot?

# data transformation


library(nycflights13)
library(tidyverse)

df <- nycflights13::flights
ggplot(data = df, mapping = aes(x = year, y= flight, colour = origin))+
  geom_point() + facet_wrap(~origin)

summary(df)

# Filter

filter(df, month==1, day==1)

jan <- filter(df, month==1)
ggplot(jan, mapping = aes(x = dep_delay, y = carrier, colour = flight)) + 
  geom_point() + theme_classic()





feb <- filter(df, month==2)
march <- filter(df, month==3)
april <- filter(df, month==4)
may <- filter(df, month==5)
june <- filter(df, month==6)



# Checking Number of rows

nrow(jan)
nrow(feb)
nrow(april)
nrow(may)
nrow(june)

# Checking number of columns

ncol(jan)
ncol(feb)
ncol(march)
ncol(april)
ncol(may)
ncol(june)

# saving in excel 

write_xlsx(jan, "jan.xlsx")
write_xlsx(feb, "feb.xlsx")
write_xlsx(march, "march.xlsx")
write_xlsx(april, "april.xlsx")
write_xlsx(may, "may.xlsx")
write_xlsx(june, "june.xlsx")

#fetching multiple months data 

jan <- filter(df, month==1, day==1, carrier=="UA")
jan <- filter(df, month==1)


jan_feb_march <- filter(df, month %in% c(1, 2, 3))
jan_feb_march_april <- filter(df, month %in% c(1, 2, 3, 4))
jan_feb_march_april_may <- filter(df, month %in% c(1, 2, 3, 4, 5))
jan_feb_march_april_may_june <- filter(df, month %in% c(1, 2, 3, 4, 5, 6))


filter(df, arr_delay > 120 & arr_delay <300)

filter(df, arr_delay > 120:300)

```


# Day 4
```{r}

df <- nycflights13::flights


library(dplyr)
glimpse(df)  # for summary


# iris example

data("iris"
)
iris %>% 
  group_by(Species) %>%
  summarise(avg = mean(Petal.Length)) %>%
  arrange(avg)

# flights data

df %>% 
  group_by(carrier) %>% 
  summarise(add = sum(hour)) %>%  #pipe operator is used to continue a process
  arrange(add)



# Sub-setting of data

df1 <- df %>% select(year, month, day)

# You can also arrange th e columns in order just by changing the name place as 1st, 2nd etc.


df2 <- df %>% select(year, month, day) %>% filter(month == 1)

df %>% select(year:arr_time)


df %>% rename(depature_time = dep_time)



# difference between wide and long format


df %>%  arrange(arr_time)
arr_time <- df %>%  arrange(arr_time)   # Ascending Order


arr_time1 <- df %>%  arrange(desc(arr_time)) 

df %>% distinct(flight)









data("iris")


#PCA

x <- prcomp(iris[ , -5], center = TRUE, scale. = TRUE)
print(x)
summary(x)

# ggplot2 grouping

iris <- cbind(iris, x$x)

ggplot(iris, aes(PC1, PC2, colour = Species, fill = Species)) + 
  stat_ellipse(geom = "polygon", colour = "black", alpha = 0.5)+
  geom_point(shape = 21, colour = "black")


# PCA plot
# install.packages("factoextra")
# install.packages("FactoMineR")

library(factoextra)
library(FactoMineR)




# PCA table

iris.pca <- PCA(iris[ , -5], graph = TRUE,scale.unit = TRUE)



# Screeplot

fviz_eig(iris.pca, addlabels = T)
fviz_eig(iris.pca, addlabels = T, ylim = c(0, 50))

# PCA Plot

fviz_pca_var(iris.pca, col.var = "cos2",
             gradient.col = c("black", "blue", "green", "red"),
                              repel = TRUE)+
  labs(title = "PCA of Parameters", x = "PC1 (49%)", y = "PCA2 (23%)",
colour = "cos2")


ggsave("PCA.pdf", units = "in", width = 6.5, height = 5.5)
```



# Day 5

```{r}
diamonds
ggplot(data = diamonds, mapping = aes(x = carat, y = price, col = cut)) +
  geom_point()


library(dplyr)
library(ggplot2)
colnames(df)


ggplot(data = df, mapping = aes(x = hour, y = arr_delay, col = carrier)) +
  geom_point()

# Correlation Plots

# to check normal distribution 

hist(df$distance)
ggsave("hist_plot.pdf")
qqnorm(df$distance)

ggsave("qqmorm_plot.pdf")




# data generate

normal_data <- rnorm(200)
hist(normal_data, col = "lightblue")
ggsave("normal_data.pdf")

qqnorm(normal_data, col = "red")
ggsave("normal_data1.pdf")


shapiro.test(normal_data)


# to make data correct   log transformation, sequare root transformation,  cube root 
# transformation, scaling transformation, mini

# Correlation plots
# Corrplot
# install.packages("corrplot", dependencies = T)

library(corrplot)


df <- mtcars
m <- cor(df)

# specify method

m <- cor(df, method = "pearson")
?cor
corrplot(m)

corrplot(m, method = "number", order = "alphabet")

corrplot(m, method = "shade", order = "alphabet")
corrplot(m, method = "shade", order = "alphabet", type = "upper")
corrplot.mixed(m, lower = "pie", upper = "ellipse")



# clustring

corrplot(m, order = "hclust", addrect = 2)

```









# Day 7


#    ***Why we do Statistical Analysis?***
___
# ***Comparison Tests***
# **t-test** <= 30
## One Sample t-test
## Two Sample t-test
## Paired t-test / Independed t-test
## Unpaired t-test / Dependend t-test



```{r}
# Data import

df <- iris

# one sample t-test

t.test(df$Sepal.Length, mu = 6.5)

```

```{r}
# two sample t-test
## independent t-test / unpaired t-test

t.test(df$Sepal.Length, df$Petal.Length, paired = FALSE)

# Why Welch Two Sample t-test is used in independent
```
```{r}
# When to use paired t-test

pre_treatment <- c(rnorm(2000, mean = 150, sd = 10))
post_tretment <- c(rnorm(2000, mean = 144, sd = 9))

t.test(pre_treatment, post_tretment, paired = TRUE)

```

```{r}
library(tidyverse)
df <- iris
df <- df %>% 
  filter(Species %in% c("setosa", "versicolor")) 
  
t.test(df$Petal.Width ~ df$Species)
```

___
## **Z test** >= 30





___
# **ANOVA**
## If variables are greater then 2
## Different types of ANOVA

```{r}

# ANOVA
## Data importing

library(palmerpenguins)
df <- penguins

colnames(df)

# find unique values

unique(df$species)
unique(df$island)



# t.test on which column

t.test(df$bill_length_mm ~ df$sex)
library(tidyverse)
ggplot(df, aes(sex, bill_length_mm, fill = sex)) + geom_boxplot()

data <- df %>% 
  select(species, flipper_length_mm) %>% 
  drop_na()



  
# summary

summary(data)

data %>% 
  group_by(species) %>% 
  summarize(mean(flipper_length_mm))

ggplot(data) + 
  aes(species, flipper_length_mm, col = species) + 
  geom_jitter() + theme(legend.position = "none")


# normal distribution

hist(data$flipper_length_mm)

shapiro.test(data$flipper_length_mm)

# whenever to check normal distribution only check on single specie

data %>% 
  group_by(species) %>% 
  summarize(shapiro_pvalue = shapiro.test(flipper_length_mm)$p.value)

s_w <- shapiro.test(data$flipper_length_mm)

summary(s_w)

# ANOVA

res_anova <- aov(flipper_length_mm ~ species, data = data)
summary(res_anova)

# What are values represented in anova table like, Df, Sq, F value

hist(res_anova$residuals)
qqnorm(res_anova$residuals)
shapiro.test(res_anova$residuals)


# to check wether our data is homogeneous or not

library(car)
leveneTest(data$flipper_length_mm ~ species, data = data)
library(lattice)
dotplot(data$flipper_length_mm ~ species, data = data)
ggplot(data)+
  aes(species, flipper_length_mm, fill = species)+
  geom_boxplot()



ggplot(data)+
  aes(species, flipper_length_mm, fill = species)+
  geom_boxplot() + theme(legend.position = "none") # theme(legend.position) to remove legends

# outliers removal equation?


# calculate mean and sd for plots and ANOVA plots

aggregate(flipper_length_mm ~ species, data = data,
          function(x) round(c(mean = mean(x), sd = sd(x)), 2))
          

# another way 

df1 <- group_by(data, species) %>% 
  summarise(
    mean = mean(flipper_length_mm, na.rm = TRUE),
    sd = sd(flipper_length_mm, na.rm = TRUE)
  )
df1

# ANOVA Test 1

oneway.test(flipper_length_mm ~ species, data = data, var.equal = TRUE)
oneway.test(flipper_length_mm ~ species, data = data, var.equal = FALSE)

# ANOVA Test 2 Recommended Test

res_anova <- aov(flipper_length_mm ~ species, data = data)
summary(res_anova)


# Posthoc Test

library(multcomp)

# TukeyHSD

res_anova <- aov(flipper_length_mm ~ species, data = data)
summary(res_anova)
post_test <- glht(res_anova, linfct = mcp(species = "Tukey"))

summary(post_test)

plot(post_test)




plot(TukeyHSD(res_anova))


?mcp

res_anova <- aov(flipper_length_mm ~ species, data = data)
summary(res_anova)
dun_test <- glht(res_anova, linfct = mcp(species = "Dunnet"))
summary(dun_test)



# Two-way ANOVA 

library(ggthemes)
library(multcompView)
library(tidyverse)


df5 <- ToothGrowth
View(ToothGrowth)
df5$dose <- as.factor(df5$dose)

# Calculate two-way anova

anova <- aov(len ~ supp*dose, data = df5)
summary(anova)

plot(anova)


# Tukey-HSD

tukey <- TukeyHSD(anova)
tukey
plot(tukey)


# to remove lettering

letter <- multcompLetters4(anova, tukey)
letter

# Convert them into a df

letter1 <- data.frame(letter$`supp:dose`$Letters)
letter1

# mean and Sd

df_mean <- df5 %>% 
  group_by(supp, dose) %>% 
  summarise(len_mean=mean(len), sd=sd(len)) %>% 
  arrange(desc(len_mean))


tibble(df_mean)


# Adding letter

df_mean$letter1 <- letter1$letter..supp.dose..Letters
df_mean


# plot

plot <- ggplot(df_mean)+
  aes(dose, len_mean, group = supp, fill = supp) + 
  geom_bar(stat = 'identity', position = position_dodge(0.9)) + 
 
   geom_errorbar(aes(ymin = len_mean-sd, ymax = len_mean+sd), 
                width = 0.1, position = position_dodge(0.9)) + 
 
   geom_text(aes(label = letter1, y = len_mean+sd), vjust = -0.4,
            position = position_dodge(0.9)) + 
  scale_fill_manual(values = c("lightblue", "lightpink")) + 
  labs(x = "Dose", y = "Length (mm)", 
       title = "Tooth Growth", fill = "Supplement");plot


ggsave("plot.pdf", width = 10, height = 8, units = "in")



```



